### 2019-01-08 paper/main.tex
Retoques paper

	Softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems but not the best when working with ordinal data. In this paper, the Proportional Odds Model (POM) will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimisation algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust results.

@@

	\section{Related work}
	\label{sect:relatedwork}
	J. de la Torre et al. proposed the use of QWK loss function for the optimisation algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces the overfitting and the required training time. Also, they checked the importance of hyper-parameter tuning.

@@

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try and predict from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras, illumination conditions and resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition that was hosted on the Kaggle platform. Also, this dataset was used in later works \cite{de2018weighted}\cite{nebot2016diabetic} and ordinal techniques (such as an ordinal cost function) were applied in order to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 256 by 256 pixels. Data augmentation techniques are applied to achieve a higher number of samples.

		\item \textit{Adience}\footnote{http://www.openu.ac.il/home/hassner/Adience/data.html}. This dataset consists of $26580$ faces belonging to $2284$ subjects. We use the form of the dataset where faces have been pre-cropped and aligned. The dataset was preprocessed, using the methods described in a previous work \cite{beckham2017unimodal}, so that the images are 256px in width and height, and pixels values follow a $(0;1)$ normal distribution. The original dataset is split into five cross-validation folds. The training set consists of merging the first four folds which comprise a total of $15554$ images. From this, $10\%$ of the images are held out as part of a validation set. The last fold is used as test set.
	\end{itemize}

@@

	Quadratic Weighted Kappa Loss, that J. de la Torre described in previous work, is considered as loss function for this optimizer as it gives better performance for ordinal classification problems.

@@

@@

@@

	\subsection{Statistical analysis}
	The significance and relative importance of the parameters concerning the results obtained, as well as suitable values for each, were obtained using ANalysis Of the VAriance (ANOVA)

	The ANalysis Of the VAriance (ANOVA) \cite{miller1997beyond} is one of the most widely used statistical techniques. ANOVA is essentially a method of analysing the variance to which a response is subject into its various components, corresponding to the sources of variation which can be identified.

	ANOVA, in this case, examines the effects of three quantitative variables (termed factors) on one quantitative response. Considered factors are the link function, which can be logit, probit and complementary log-log, the learning rate for the Adam optimisation algorithm ($10^{-2}$, $10^{-3}$ and $10^{-4}$) and the batch size (5, 10 and 15).

	Following the setup of the previous study, we performed an ANOVA III analysis and multiple comparison tests. We assume that five executions are enough to do the statistical tests because of the computational time limitations. The test shows that there is no batch size whose results are significatively better than the results of all other batch sizes. This does not mean that these differences could not exist for specific numbers of samples per batch. So, in order to determine for each type of function whether a batch size is better than the others, we have performed an ANOVA I analysis - where the only factor is the batch size - and multiple comparison tests.

@@

	We consider some hypotheses testing where the null hypothesis is proposed that each term of the above equation is independent of the levels involved. The hypotheses for the levels the L factor are
	\[
		H_0 \equiv L_1 = L_2 = L_3
	\]
	\[
		H_1 \equiv \text{ some } L_i \text{ is different}
	\]

@@

@@

>>>>

	Softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems but not the best when working with ordinal data. In this paper, the Proportional Odds Model (POM) will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimization algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust results.

@@

	\section{Related work}
	\label{sect:relatedwork}
	J. de la Torre et al. proposed the use of QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces the overfitting and the required training time. Also, they checked the importance of hyper-parameter tuning.

@@

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try and predict from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras, illumination conditions and resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition that was hosted on the Kaggle platform. Also, this dataset was used in later works \cite{de2018weighted}\cite{nebot2016diabetic} and ordinal techniques (such as an ordinal cost function) were applied in order to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels. Data augmentation techniques are applied to achieve a higher number of samples.

		\item \textit{Adience}\footnote{http://www.openu.ac.il/home/hassner/Adience/data.html}. This dataset consists of $26580$ faces belonging to $2284$ subjects. We use the form of the dataset where faces have been pre-cropped and aligned. The dataset was preprocessed, using the methods described in a previous work \cite{beckham2017unimodal}, so that the images are 256 pixels in width and height, and pixels values follow a $(0;1)$ normal distribution. The original dataset is split into five cross-validation folds. The training set consists of merging the first four folds which comprise a total of $15554$ images. From this, $10\%$ of the images are held out as part of a validation set. The last fold is used as test set.
	\end{itemize}

@@

	Quadratic Weighted Kappa Loss is considered as loss function for this optimizer as it gives better performance for ordinal classification problems.

@@

	 Experiments were run with the standard cross-entropy loss and the softmax function too in order to prove the performance improvement of the QWK loss and the Proportional Odds Model. The results of these experiments are analysed in Appendix \ref{app:NominalComparison}.

@@

	The best mean WK value was obtained with the complementary log-log link function using a batch size of 10 and a learning rate of $10^{-3}$.

@@

	The best mean WK value was obtained with the probit link function using a batch size of 15 and a learning rate of $10^{-2}$.

	\subsection{Statistical analysis}
	In this subsection, a statistical analysis will be performed in order to obtain robust conclusions from the results presented in this section.

	The significance and relative importance of the parameters concerning the results obtained, as well as suitable values for each, were obtained using the ANalysis Of the VAriance (ANOVA) test.

	The ANOVA test \cite{miller1997beyond} is one of the most widely used statistical techniques. ANOVA is essentially a method of analysing the variance to which a response is subject into its various components, corresponding to the sources of variation which can be identified.

	ANOVA, in this case, examines the effects of three quantitative variables (termed factors) on one quantitative response. Considered factors are the link function, which can be logit, probit and complementary log-log, the learning rate for the Adam optimization algorithm ($10^{-2}$, $10^{-3}$ and $10^{-4}$) and the batch size (5, 10 and 15).

	Following the setup of the previous study, we performed an ANOVA III analysis and multiple comparison tests. We assume that five executions are enough to do the statistical tests because of the computational time limitations.

	The test shows that there is no batch size whose results are significatively better than the results of all other batch sizes. This does not mean that these differences could not exist for specific numbers of samples per batch. So, in order to determine for each type of function whether a batch size is better than the others, we have performed an ANOVA I analysis - where the only factor is the batch size - and multiple comparison tests.

@@

	We consider some hypotheses testing where the null hypothesis is proposed that each term of the above equation is independent of the levels involved. The hypotheses for the levels of the L factor are
	\[
		H_0 \equiv L_1 = L_2 = L_3
	\]
	\[
		H_1 \equiv \text{ some } L_i \text{ is different}
	\]

@@

	The proposed POM model has improved the performance of the deep network regarding Weighted Kappa compared to the model that uses the softmax function. This enhancement is more notable for the Diabetic Retinopathy dataset. Also, it reduces the chance that the model gets stuck when training with some parameter configurations. So, the most significant improvements of these link functions are the WK performance increase, the reduction of the number of parameters configurations that should be tried to find the best one and the prevention of the over-fitting and the stagnation.

@@

	\appendix
	\section{Statistical comparison between nominal and ordinal methods}
	\label{app:NominalComparison}

	The same experiments described in section \ref{sect:experiments} were repeated using the cross-entropy instead of the QWK as loss function for the optimiser and the softmax function instead of the Proportional Odds Model for the output of the network.
	The results for both datasets are shown in Table \ref{table:NominalResults}.

	\begin{table}[ht]
		\footnotesize
		\centering
		\begin{tabular}{lcc|cc}
			Dataset & BS & LR & $\kappa_{val}$ & $\kappa_{test}$\\\hline\addlinespace[0.05cm]
			Retinopathy & $5$ & $10^{-2}$ & $0.2210$ & $0.2050$\\
			Retinopathy & $5$ & $10^{-3}$ & $0.2807$ & $0.2654$\\
			Retinopathy & $5$ & $10^{-4}$ & $0.4580$ & $0.4449$\\

			Retinopathy & $10$ & $10^{-2}$ & $0.2972$ & $0.2997$\\
			Retinopathy & $10$ & $10^{-3}$ & $0.3972$ & $0.3989$\\
			Retinopathy & $10$ & $10^{-4}$ & $\textbf{0.4972}$ & $\textbf{0.4967}$\\

			Retinopathy & $15$ & $10^{-2}$ & $0.3692$ & $0.3676$\\
			Retinopathy & $15$ & $10^{-3}$ & $0.4107$ & $0.4162$\\
			Retinopathy & $15$ & $10^{-4}$ & $0.4929$ & $0.4859$\\
			\hline\addlinespace[0.05cm]
			Adience & $5$ & $10^{-2}$ & $0.0605$ & $0.0769$\\
			Adience & $5$ & $10^{-3}$ & $0.7994$ & $0.7313$\\
			Adience & $5$ & $10^{-4}$ & $0.8009$ & $0.7383$\\

			Adience & $10$ & $10^{-2}$ & $0.0559$ & $0.0576$\\
			Adience & $10$ & $10^{-3}$ & $0.8239$ & $0.7436$\\
			Adience & $10$ & $10^{-4}$ & $0.7938$ & $0.7218$\\

			Adience & $15$ & $10^{-2}$ & $0.0593$ & $0.0713$\\
			Adience & $15$ & $10^{-3}$ & $\textbf{0.8309}$ & $\textbf{0.7575}$\\
			Adience & $15$ & $10^{-4}$ & $0.7867$ & $0.7129$\\
		\end{tabular}
		\caption{Adience results. BS stands for Batch Size, LR for Learning Rate and LF for link function.}
		\label{table:NominalResults}
	\end{table}

	As we did in Section X, we must analyse the effect of the batch size (with values 5, 10 and 15) and the learning rate ($10^{-2}$, $10^{-3}$ and $10^{-4}$) factors have over the WK metric. So, we make an ANOVA II analysis for each dataset.

	In the Diabetic Retinopathy dataset, there are significative differences between the mean values of WK depending on the batch size and the initial value of the learning rate. Also, no significative interaction was detected between these factors.

	After that, a post-hoc Tukey's test shows that the best value for the batch size factor is 15 followed by 10. However, there are no significative differences between them, but there are differences with size 5. Also, the best value for the learning rate is $10^{-4}$ and there are significative differences with the rest of values.

	A similar study was made for the Adience dataset. There are not significative differences for the mean value of WK w.r.t. the batch size factor. Also, there are no interactions between the batch size and the learning rate. The best results were obtained with $10^{-3}$ for the learning rate followed by $10^{-4}$, without significative differences between them.

	Finally, the conclusions of this study are exposed below:
	\begin{itemize}
		\item There are no interactions between both factors.
		\item Batch sizes 10 and 15 obtained the best results.
		\item Learning rate $10^{-4}$ is the best choice for both datasets as it the best for Diabetic Retinopathy, with significative differences, and it is the second for Adience, without significative differences with the best. 
	\end{itemize}
